### 2022-03-16 paper/localpsf.tex
restructuring paper

The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$. H-matrices are a matrix format in which the matrix is re-ordered then recursively subdivided into blocks, so that many off-diagonal blocks of the matrix are low rank. For more details about H-matrices, see Appendix \ref{app:h_matrix}. While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply. Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems. The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.

@@

In this section we present the three main concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Third, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

@@

Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to construct an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$. This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

>>>>

The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$. H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank. For more details about H-matrices, see Appendix \ref{app:h_matrix}. While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply. Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems. The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.

@@

In this section we present the three main concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

@@

Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$. This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.
